\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{graphicx}

\author{Maarten de Jonge}
\date{\today}
\title{Concurrency and Parallel Programming \\
       \large{Assignment 4 - Hadoop and MapReduce}}

\begin{document}
\maketitle

\section*{Hadoop}

\subsection*{Purpose}
Hadoop is a framework for processing huge datasets on huge clusters of
commodity hardware. When used on large sets of independent data, it's relatively
easy to spread the workload over tens to thousands of machines, leading to
immense scalability.

\subsection*{Architecture}
Hadoop consists mainly of two parts; a distributed filesystem (HDFS) and a MapReduce
engine.

\subsubsection*{HDFS}
HDFS consists of a single \emph{namenode} along with a cluster of
\emph{datanodes}. The namenode contains the directory tree and locations of each
file within the cluster, while the datanodes hold the actual data. The data can
be stored with a configurable level of redundancy, and failures of the datanodes
will generally not be much of a problem. However, failure of the namenode will
render the cluster unreachable.

\subsubsection*{MapReduce}
The MapReduce engine consists of a \emph{jobtracker}, a number of
\emph{tasktrackers} and the \emph{Map} and \emph{Reduce} functions.

The jobtracker is responsible for dividing the workload between the available
tasktrackers.

The \texttt{Map} function takes a list of key-value pairs, and produces a new
list of key-value pairs. The resulting data is grouped by key, and the
\texttt{Reduce} function receives one key and a list of values corresponding to
that key (according to the output of \texttt{Map}), returning a single value per
key.

\section*{The Biojava Stuff}
Not finished yet!

\end{document}
